---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  error = TRUE,
  message = FALSE
)
```

# matSPACE

<!-- badges: start -->
[![R-CMD-check](https://github.com/kimhyew1/matSPACE/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/kimhyew1/matSPACE/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

This is a README file of the R package _matSPACE_. Our package fits sparse partial correlation networks for matrix-variate data, i.e. observations that are themselves p x q matrices rather than a single vector of q measurements. We extend the SPACE (Sparse PArtial Correlation Estimation) joint estimation framework to a Kronecker-product precision structure, `Omega = V %x% U`, where `V` (q x q) is the column-wise precision matrix and `U` (p x p) is the row-wise precision matrix. All partial correlations are estimated jointly via an L1-penalized (lasso) shooting algorithm, which preserves symmetry of the estimated network and avoids the tuning-parameter selection difficulties of separate node-wise regressions.

## Installation of the package

To install our package, you may simply execute the following codes:
```{r, eval = FALSE}
install.packages("matSPACE")

# development version
# install.packages("pak")
pak::pak("kimhyew1/matSPACE")
```

## A basic example of using the package

We give a toy example to apply the main function `matSPACE`, which estimates the Kronecker-structured row/column precision matrices `U` and `V`.

### Generate data

We first generate simulated matrix-variate data. Columns 1-2 and columns
3-4 are correlated by construction, so there is real sparse structure
for `matSPACE` to recover (pure noise data has no structure to
penalize away, so the BIC-optimal fit is always the fully sparse one).
```{r, eval = TRUE}
set.seed(1)
p = 5
q = 4
n = 20

library(matSPACE)
data = replicate(n, {mat = matrix(rnorm(p * q), p, q)
                     mat[, 2] = mat[, 2] + 0.8 * mat[, 1]
                     mat[, 4] = mat[, 4] + 0.8 * mat[, 3]
                     mat}, simplify = FALSE)
```

### Model fitting

Applying `matSPACE` to the simulated data, a lasso penalty path is searched for `V` and `U` separately, and the BIC-minimizing fit is returned for each.
```{r, eval = TRUE}
fit = matSPACE(data, K = 10)
fit$V   # q x q column-wise precision matrix
fit$U   # p x p row-wise precision matrix
```

### Visualization of the lambda path

The full lambda path behind this selection is attached as a `"path"` attribute, which lets us plot BIC against lambda without refitting.
```{r, eval = TRUE}
path = attr(fit, "path")
plot(log(path$V$lambda), path$V$bic, type = "b",
     xlab = "log(lambda)", ylab = "BIC", main = "Column (V) lambda path")
i_min = which.min(path$V$bic)
points(log(path$V$lambda[i_min]), path$V$bic[i_min], col = "red", pch = 19)
```

### Fitting with a single lasso penalty

If you already know the penalty you want, `space` fits a single lasso penalty directly, without searching a path.
```{r, eval = TRUE}
fit_single = space(data, lam = 0.5)
fit_single$ParCor
```

## Issues

We are happy to troubleshoot any issue with the package;

- please contact to the maintainer by kimhw4126@gmail.com, or
- please open an issue in the github repository.
